---
title: "Practical machine learning course project"
author: "Joaquim Villen Benseny"
date: "19 August 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
The following paper aims to predict a classification of excercises based on the utilisation of different accelerometers for different users. 

## Data exploration

### Data reading
```{r}
training <- read.csv('pml-training.csv', na.strings = c("NA", "#DIV/0!", ""))
crossval <- read.csv('pml-testing.csv', na.strings = c("NA", "#DIV/0!", ""))
```

### Data cleaning
```{r}
library(caret)
#remove entries with null values
training<- training[, colSums(is.na(training)) == 0]
crossval <- crossval[, colSums(is.na(crossval)) == 0] 

#keep output
classe <- training$classe

#remove non relevant variables
training<- training[, !(grepl("^X|timestamp|window", names(training)))]
#remove non numeric variables
training<- training[, sapply(training, is.numeric)]
training$classe<- classe

#splitting data in train test split
inTrain <- createDataPartition(training$classe, p=0.60, list=F)
training <- training[inTrain, ]
testing <- training[-inTrain, ]

#keep vector of predictors
predictors <- colnames(testing)
predictors <- predictors[!(predictors %in% c("classe"))]
```

## Data modeling

A couple of different machine learning algorithms will be trained below (random forest and gradient boosting).
```{r}
library(rpart)
mdl1 <- train(classe~.,data = training, method = 'rf', ntree = 10)
mdl1
```

Showing the tree:
```{r}
library(rpart)
library(rpart.plot)
rtree<- rpart(classe ~ ., data=training, method="class")
prp(rtree)
```


## Results and conclusions
Test error
```{r}
pred <- predict(mdl1, testing)
confusionMatrix(pred,testing$classe)$overall
```

```{r}
error <- 1 - as.numeric(confusionMatrix(testing$classe, pred)$overall[1])
```

The results predicted for the test dataset are provided below:
```{r}
predict(mdl1,crossval[,predictors])
```

Therefore it is possible to conclude that the out of sample error of the random forest model trained above is around `r error`. The  model accuracy is around `r confusionMatrix(pred,testing$classe)$overall["Accuracy"]*100` 